Miniproject - Introduction to Schemes

### 2022-03-26 Schemes/submission.tex
Finalized schemes

    To show uniqueness, assume there are $\alpha/f^N, \beta/f^N \in \D{f}$ with
    \begin{equation*}
        \restr{\frac {\alpha} {f^N}}{\D{g_if}} = \restr{\frac {\beta} {f^N}}{\D{g_if}} \quad \text{for all $i$}
    \end{equation*}
    Therefore there is $n_i \in \N$ such that
    \begin{equation*}
        (fg_i)^{n_i} (\alpha f^N + \beta f^N) = 0
    \end{equation*}

@@

    Since $\bigcup_i \D{(fg_i)^N} = \bigcup_i \D{fg_i} = \D{f}$ it follows that $f^n \in \langle (fg_i)^N \ | \ i \rangle$ for some $n \in \N$ (again the Covering Trick, Lemma~\ref{prop:covering_trick}).
    Therefore we find
    \begin{equation*}
        f^n(\alpha - \beta) \in \langle (fg_i)^N \ | \ i \rangle (\alpha - \beta) = \langle (fg_i)^N (\alpha - \beta) \ | \ i \rangle = \{ 0 \}
    \end{equation*}
    and so $\alpha = \beta \in S_f$.

    Now we show existence.
    By the uniqueness above, it follows that
    \begin{equation*}
        \restr{s_i}{\D{fg_ig_j}} = \restr{s_i}{\D{fg_i} \cap \D{fg_j}} = \restr{s_j}{\D{fg_i} \cap \D{fg_j}} = \restr{s_j}{\D{fg_ig_j}}
    \end{equation*}
    wlog have again a finite cover, i.e. only finitely many $g_i$.
    Hence find an $N \in \N$ such that each $s_i = s'_i / (fg_i)^N$ with $s'_i \in S$ homogeneous.
    By possibly replacing $N$ with a bigger $N$, we can now assume that
    \begin{equation*}
        (f^2g_ig_j)^N \left( s'_i (fg_j)^N - s'_j (fg_i)^N \right) = 0 \quad \text{as $\restr{s_i}{\D{fg_ig_j}} = \restr{s_j}{\D{fg_ig_j}}$}
    \end{equation*}
    Now note that
    \begin{equation*}
        s_i = \frac {a_i} {b_i} \quad \text{with}\ a_i = s'_i(fg_i)^N, \ b_i = (fg_i)^{2N}
    \end{equation*}
    and
    \begin{equation*}
        a_i b_j - a_j b_i = s'_i (fg_i)^N (fg_j)^{2N} - s'_j (fg_j)^N (fg_i)^{2N} = \underbrace{(f^2g_ig_j)^N \left(s'_i (fg_j)^N - s'_j (fg_i)^N \right)}_{= 0}
    \end{equation*}
    Now observe that $\D{b_i} = \D{fg_i}$ and so $f^n \in \langle b_i \ | \ i \rangle$ for some $n$ (once again the Covering Trick, Lemma~\ref{prop:covering_trick}).
    Let $f^n = \sum_i r_i b_i$ and get
    \begin{equation*}
        a_i f^n = \sum_l r_l b_l a_i = \sum_l r_l a_l b_i = b_i \sum_l r_l a_l
    \end{equation*}
    Note that $a_i, b_i, f$ are homogeneous, and so we can also choose $r_i$ to be homogeneous.
    Then find that $0 = \deg(s_i) = \deg(a_i) - \deg(b_i) = \deg(\sum r_l a_l) - n\deg(f)$.

@@

    Now consider the natural transformation
    \begin{equation*}
        \eta: \O_{\Spec R} \Rightarrow \phi_*\Bigl( \restr{\O_{\Proj(S)}}{\D{f}} \Bigr)
    \end{equation*}
    given on basic open sets by
    \begin{equation*}
        \eta_{\D{g/f^n}}: R_{g/f^n} \to (S_{fg})_0, \quad \frac {h/f^m} {(g/f^n)^l} \mapsto \frac {h f^{nl}} {g^l f^m}
    \end{equation*}
    Clearly this is a ring isomorphism, so $\eta$ is a natural isomorphism.
    The claim follows.
\end{proof}
\begin{corollary}
    $(\Proj(S), \O_{\Proj(S)})$ is a scheme.
\end{corollary}
\begin{definition}
    For any ring $R$ and $n \in \N$, define the \emph{projective $n$-space} as\footnote{Endow $R[x_0, ..., x_n]$ with the standard natural grading, i.e. $\deg(x_i) = 1$. Note that this makes $R[x_0, ..., x_n]_0 = R$, so all $r \in R$ are homogeneous of degree $0$.}
    \begin{equation*}
        \P_R^n := \Proj(R[x_0, ..., x_n])
    \end{equation*}
    which naturally becomes a scheme over $\Spec(R)$ via the morphism given by
    \begin{align*}
        &f: \Proj(R[x_0, ..., x_n]) \to \Spec(R), \quad \p \mapsto \p \cap R \\
        &f^\#_{D_f}: R_f \to \O_{\Proj(R[x_0, ..., x_n])}(\D{f}) = R_f[x_0, ..., x_n], \quad r \mapsto r
    \end{align*}
\end{definition}

\section{Projective space as a variety}
For this section, let $R$ be a ring and $n \in \N$.
\begin{prop}
    The morphism $\P_R^n \to \Spec(R)$ is separated.
\end{prop}
\begin{proof}
    This is the same proof as for \citestacks{01MC}.
    By Lemma~\ref{prop:Df_affine} we find a cover $\Proj(S) = \bigcup_i \D{x_i}$ of affine opens.
    We use the characterization from the lecture and show that $\D{x_i} \cap \D{x_j}$ is an affine open and the canonical multiplication map
    \begin{equation*}
        \O_{\P_R^n}(\D{x_i}) \times \O_{\P_R^n}(\D{x_j}) \to \O_{\P_R^n}(\D{x_i} \cap \D{x_j})
    \end{equation*}
    is surjective.

@@

    However, for $f/(x_ix_j)^n \in (R[x_0, ..., x_n]_{x_ix_j})_0$ we have that
    \begin{equation*}
        \frac f {x_i^{2n}} \cdot \frac {x_i^n} {x_j^n} = \frac f {(x_ix_j)^n}
    \end{equation*}
    and clearly $f/x_i^{2n} \in (R[x_0, ..., x_n]_{x_i})_0$ as $\deg(f) = \deg(x_i) n + \deg(x_j) n = 2n$ and $x_i^n/x_j^n \in (R[x_0, ..., x_n]_{x_j})_0$.
    The claim follows.
\end{proof}
To study integrality, it is useful to have the next small lemma on the nilradical.
\begin{lemma}
    \label{prop:nil_poly_ring}
    If $\Spec(R)$ is irreducible, then $\Nil(R)R[x_0, ..., x_n] = \Nil(R[x_0, ..., x_n])$.
\end{lemma}
\begin{proof}
    Clearly $\Nil(R)R[x_0, ..., x_n] \subseteq \Nil(R[x_0, ..., x_n])$, as for $\sum_{i = 1}^m f_i r_i$ with $f_i^n = 0$ have that
    \begin{equation*}
        \Bigl( \sum_{i = 1}^m f_i r_i \Bigr)^{m n} = \sum_{1 \leq i_1, ..., i_{mn} \leq m} \underbrace{f_{i_1} ... f_{i_{mn}}}_{\mathclap{\substack{= 0 \\ \text{as at least one $i$ repeats at least $n$ times in $i_1, ..., i_{mn}$}}}} r_{i_1} ... r_{i_mn}
    \end{equation*}
    Now note that $\Nil(R[x_0, ..., x_n]) \subseteq \a$ for all prime ideals $\a$, and the claim follows since $\Nil(R)$ is prime by assumption that $\Spec(R)$ is irreducible.
\end{proof}
\begin{prop}
    Assume $\Spec(R)$ is irreducible.
    Then $\P_R^n$ is an irreducible scheme.
\end{prop}
\begin{proof}
    We show that $\D{x_0} \subseteq \P_R^n$ is dense.
    The claim then follows, as any two nonempty, disjoint open sets give nonempty, disjoint open sets in $D_{x_0}$ which is isomorphic to the 
    irreducible\footnote{This is obviously irreducible, as $R[x_0, ..., x_n]/\Nil(R[x_0, ..., x_n]) = (R/\Nil(R))[x_0, ..., x_n]$ is integral by assumption and Lemma~\ref{prop:nil_poly_ring}.}
    affine scheme $\Spec(R[x_0, ..., x_n]_{x_0})_0$ by Lemma~\ref{prop:Df_affine}.

@@

    wlog $S$ is noetherian, otherwise continue with the ring
    \begin{equation*}
        \tilde{S} := \bigl( \Z / \mathrm{char}(S) \Z \bigr) \bigl[ A/\alpha, x \bigr]
    \end{equation*}
    generated by the coefficients of $A/\alpha$ and $w$ (this is noetherian, as it is a quotient of a polynomial ring in finitely many variables).

    Now equip $S_{\mathfrak{l}}$ with the $\mathfrak{l}$-adic topology.
    Note that $S$ is noetherian and local, thus the Krull intersection theorem \cite[8.41]{comalg_notes} shows that the $\mathfrak{l}$-adic topology is Hausdorff and we find
    \begin{equation*}
        w = \lim_{i \to \infty} w = \lim_{i \to \infty} \frac {A^i} {\alpha^i} w = \Bigl( \lim_{i \to \infty} \frac {A^i} {\alpha^i} \Bigr) w = 0 w = 0
    \end{equation*}
    since $(A/\alpha)^i$ has coefficients in $\mathfrak{l}^i$, thus converges to $0$ as $i \to \infty$.

@@

But now enough of talking, here is the main result.
\begin{corollary}
    The morphism $\P_R^n \to \Spec(R)$ is universally closed.
\end{corollary}
\begin{proof}
    Use Lemma~\ref{prop:universally_closed_affines}, so consider an affine base change $f_{\Spec(S)}: \P_R^n \times_{\Spec(R)} \Spec(S) \to \Spec(S)$.
    By Lemma~\ref{prop:projective_space_base_change}, have that $\P_R^n \times_{\Spec(R)} \Spec(S) \cong \P_S^n$ as schemes over $S$, hence $f_{\Spec(S)}$ is isomorphic to $\P_S^n \to \Spec(S)$.
    This morphism is closed by Proposition~\ref{prop:projective_structure_morphism_closed} and the claim follows.
\end{proof}
\begin{corollary}
    Projective space $\P_R^n$ over $\Spec(R)$ is proper.
\end{corollary}
\begin{corollary}
    Let $k$ be an algebraically closed field.
    Then projective space $\P_k^n$ over $\Spec(k)$ is a complete variety.
\end{corollary}

\section{Valuative criterion of properness}
\label{sec:valuative_criterion_properness}
Instead of using some magical ad-hoc argument as before, there is also the following nice characterization of the notion of properness.
We will omit the proof here, but present the statement as it can give more geometric intuition on properness and why $\P_R^n$ is proper.
\begin{prop}
    \label{prop:valuative_criterion_properness}
    Let $X$ be noetherian and $f: X \to Y$ a morphism of finite type.
    The following are equivalent
    \begin{itemize}
        \item $f$ is proper
        \item For every valuation ring $R$ with field of fractions $K = \Frac(R)$ and all morphisms $\Spec(R) \to Y$, $\Spec(K) \to X$ that make the diagram
        \begin{center}
            \begin{tikzpicture}
                \node (K) at (0, 0) {$\Spec(K)$};
                \node (R) at (0, -1.5) {$\Spec(R)$};
                \node (X) at (3, 0) {$X$};
                \node (Y) at (3, -1.5) {$Y$};

                \draw [->] (K) -- (R);
                \draw [->] (K) -- (X);
                \draw [->] (R) -- (Y);
                \draw [->] (X) -- (Y) node [midway, right] {$f$};
            \end{tikzpicture}
        \end{center}
        commute, there exists a unique compatible morphism $\Spec(R) \to X$.
    \end{itemize}
\end{prop}
\begin{proof}
    Omitted, see e.g. \cite[II.4.7]{hartshorne} or \citestacks{01NH}.
\end{proof}
\begin{prop}
    \label{prop:projective_space_proper_alternative_proof}
    The morphism $\P_R^n \to \Spec(R)$ is proper.
\end{prop}
\begin{proof}[Alternative Proof]
    Since properness is preserved under base change, by Lemma~\ref{prop:projective_space_base_change} it suffices to show the statement for $\P_\Z^n \to \Spec(\Z)$.

@@

            \draw [->] (K) -- (R);
            \draw [->] (K) -- (X);
            \draw [->] (R) -- (Y);
            \draw [->] (X) -- (Y) node [midway, right] {$f$};
        \end{tikzpicture}
    \end{center}
    commutes.
    Note that $\Spec(K) = \{ \langle 0 \rangle \}$, so let $\xi_1 \in X$ be the image of $\langle 0 \rangle$ under $\Spec(K) \to X$.
    Since $\langle x_0, ..., x_n \rangle \not\subseteq \xi$, there is $x_i \not\in \xi$.
    Now let
    \begin{equation*}
        f_j := \begin{cases}
            x_j & \text{if $x_j \notin \xi_1$} \\
            x_j - x_i & \text{if $x_j \in \xi_1$}
        \end{cases}
    \end{equation*}
    Thus all $f_j$ are homogeneous of degree 1 with $\langle f_j \ | \ j \rangle = \langle x_0, ..., x_n \rangle$.
    It follows that $X = \bigcup_j \D{f_j}$ is a cover by affine opens such that $\xi_1 \in \bigcap_j \D{f_j}$.
    Now possibly apply a linear automorphism $f_j \to x_j$ and thus assume wlog that $f_j = x_j$.
    In particular, find $x_j \in \O_{X, \xi_1}^*$.

    The morphism $\Spec(K) \to X$ now gives an inclusion $k(\xi_1) \subseteq K$.
    Since $x_j \in \O_{X, \xi_1}^*$, observe that $g_{ij} := x_i / x_j \in K$ is well-defined and nonzero with $g_{ij} g_{jk} = g_{ik}$.

    As $R$ is a valuation ring, we find $k$ such that $g_{k0}$ is the smallest element (w.r.t. divisibility by $R$) among $g_{00}, ..., g_{n0}$.
    So we see that $g_{ik} = g_{i0} / g_{k0} \in R$ for all $i$.
    In particular, we find a ring homomorphism
    \begin{equation*}
        \Z\left[ \frac {x_0} {x_k}, ..., \frac {x_n} {x_k} \right] \to R, \quad \frac {x_j} {x_k} \mapsto g_{ik}
    \end{equation*}
    that is compatible with the field inclusion $k(\xi_1) \subseteq K$.

@@

    Also it is not too hard to show that the morphism is unique, and the claim follows.
\end{proof}
Note that in some way, this is the same proof as before.
The core idea - namely the choice of the smallest element $g_{k0}$ (w.r.t. divisibility by $R$) - more or less achieves the same effect as the more complicated ideal argument from Lemma~\ref{prop:adding_coeff_ideal_does_not_introduce_irrelevant_ideal}.
\printbibliography
\end{document}

>>>>

    To show uniqueness, assume there are $\alpha/f^N, \beta/f^N \in \D{f}$ with
    \begin{equation*}
        \restr{\frac {\alpha} {f^N}}{\D{g_if}} = \restr{\frac {\beta} {f^N}}{\D{g_if}} \quad \text{for all $i$}
    \end{equation*}
    Therefore there is $n_i \in \N$ such that
    \begin{equation*}
        (f g_i)^{n_i} (\alpha f^N + \beta f^N) = 0
    \end{equation*}

@@

    Since $\bigcup_i \D{(fg_i)^N} = \bigcup_i \D{fg_i} = \D{f}$ it follows that $f^n \in \langle (fg_i)^N \ | \ i \rangle$ for some $n \in \N$ (again the Covering Trick, Lemma~\ref{prop:covering_trick}).
    Therefore we find
    \begin{equation*}
        f^n(\alpha - \beta) \in \langle (f g_i)^N \ | \ i \rangle (\alpha - \beta) = \langle (f g_i)^N (\alpha - \beta) \ | \ i \rangle = \{ 0 \}
    \end{equation*}
    and so $\alpha = \beta \in S_f$.

    Now we show existence.
    By the uniqueness above, it follows that
    \begin{equation*}
        \restr{s_i}{\D{fg_ig_j}} = \restr{s_i}{\D{fg_i} \cap \D{fg_j}} = \restr{s_j}{\D{fg_i} \cap \D{fg_j}} = \restr{s_j}{\D{fg_ig_j}}
    \end{equation*}
    wlog have again a finite cover, i.e. only finitely many $g_i$.
    Hence find an $N \in \N$ such that each $s_i = s'_i / (fg_i)^N$ with $s'_i \in S$ homogeneous.
    By possibly replacing $N$ with a bigger $N$, we can now assume that
    \begin{equation*}
        (f^2g_ig_j)^N \left( s'_i (f g_j)^N - s'_j (f g_i)^N \right) = 0 \quad \text{as $\restr{s_i}{\D{fg_ig_j}} = \restr{s_j}{\D{fg_ig_j}}$}
    \end{equation*}
    Now note that
    \begin{equation*}
        s_i = \frac {a_i} {b_i} \quad \text{with}\ a_i = s'_i(f g_i)^N, \ b_i = (f g_i)^{2N}
    \end{equation*}
    and
    \begin{equation*}
        a_i b_j - a_j b_i = s'_i (f g_i)^N (f g_j)^{2N} - s'_j (f g_j)^N (f g_i)^{2N} = \underbrace{(f^2g_ig_j)^N \left(s'_i (fg_j)^N - s'_j (fg_i)^N \right)}_{= 0}
    \end{equation*}
    Now observe that $\D{b_i} = \D{fg_i}$ and so $f^n \in \langle b_i \ | \ i \rangle$ for some $n$ (once again the Covering Trick, Lemma~\ref{prop:covering_trick}).
    Let $f^n = \sum_i r_i b_i$ and get
    \begin{equation*}
        a_i f^n = \sum_l r_l b_l a_i = \sum_l r_l a_l b_i = b_i \sum_l r_l a_l
    \end{equation*}
    Note that $a_i, b_i, f$ are homogeneous, and so we can also choose $r_i$ to be homogeneous.
    Then find that $0 = \deg(s_i) = \deg(a_i) - \deg(b_i) = \deg(\sum r_l a_l) - n\deg(f)$.

@@

    Now consider the natural transformation
    \begin{equation*}
        \eta: \O_{\Spec R} \Rightarrow \phi_*\Bigl( \restr{\O_{\Proj(S)}}{\D{f}} \Bigr)
    \end{equation*}
    given on basic open sets by
    \begin{equation*}
        \eta_{\D{g/f^n}}: R_{g/f^n} \to (S_{f g})_0, \quad \frac {h/f^m} {(g/f^n)^l} \mapsto \frac {h f^{nl}} {g^l f^m}
    \end{equation*}
    Clearly this is a ring isomorphism, so $\eta$ is a natural isomorphism.
    The claim follows.
\end{proof}
\begin{corollary}
    $(\Proj(S), \O_{\Proj(S)})$ is a scheme.
\end{corollary}
\begin{definition}
    For any ring $R$ and $n \in \N$, define the \emph{projective $n$-space} as\footnote{Endow $R[x_0, ..., x_n]$ with the standard natural grading, i.e. $\deg(x_i) = 1$. Note that this makes $R[x_0, ..., x_n]_0 = R$, so all $r \in R$ are homogeneous of degree $0$.}
    \begin{equation*}
        \P_R^n := \Proj(R[x_0, ..., x_n])
    \end{equation*}
    which naturally becomes a scheme over $\Spec(R)$ via the morphism given by
    \begin{align*}
        &f: \Proj(R[x_0, ..., x_n]) \to \Spec(R), \quad \p \mapsto \p \cap R \\
        &f^\#_{D_f}: R_f \to \O_{\Proj(R[x_0, ..., x_n])}(\D{f}) = R_f[x_0, ..., x_n], \quad r \mapsto r
    \end{align*}
\end{definition}

\section{Projective space as a variety}
For this section, let $R$ be a ring and $n \in \N$.
\begin{prop}
    \label{prop:projective_space_separated}
    The morphism $\P_R^n \to \Spec(R)$ is separated.
\end{prop}
\begin{proof}
    This is the same proof as for \citestacks{01MC}.
    By Lemma~\ref{prop:Df_affine} we find a cover $\Proj(S) = \bigcup_i \D{x_i}$ of affine opens.
    We use the characterization from the lecture and show that $\D{x_i} \cap \D{x_j}$ is an affine open and the canonical multiplication map
    \begin{equation*}
        \O_{\P_R^n}(\D{x_i}) \times \O_{\P_R^n}(\D{x_j}) \to \O_{\P_R^n}(\D{x_i} \cap \D{x_j})
    \end{equation*}
    is surjective.

@@

    However, for $f/(x_ix_j)^n \in (R[x_0, ..., x_n]_{x_ix_j})_0$ we have that
    \begin{equation*}
        \frac f {x_i^{2n}} \cdot \frac {x_i^n} {x_j^n} = \frac f {(x_ix_j)^n}
    \end{equation*}
    and clearly $f/x_i^{2n} \in (R[x_0, ..., x_n]_{x_i})_0$ as $\deg(f) = \deg(x_i) n + \deg(x_j) n = 2n$ and $x_i^n/x_j^n \in (R[x_0, ..., x_n]_{x_j})_0$.
    The claim follows.
\end{proof}
To study integrality, it is useful to have the next small lemma on the nilradical.
\begin{lemma}
    \label{prop:nil_poly_ring}
    If $\Spec(R)$ is irreducible, then $\Nil(R)R[x_0, ..., x_n] = \Nil(R[x_0, ..., x_n])$.
\end{lemma}
\begin{proof}
    Clearly $\Nil(R)R[x_0, ..., x_n] \subseteq \Nil(R[x_0, ..., x_n])$, as for $\sum_{i = 1}^m f_i r_i$ with $f_i^n = 0$ have that
    \begin{equation*}
        \Bigl( \sum_{i = 1}^m f_i r_i \Bigr)^{m n} = \sum_{1 \leq i_1, ..., i_{mn} \leq m} \underbrace{f_{i_1} ... f_{i_{mn}}}_{\mathclap{\substack{= 0 \\ \text{as at least one $i$ repeats at least $n$ times in $i_1, ..., i_{mn}$}}}} r_{i_1} ... r_{i_{m n}}
    \end{equation*}
    Now note that $\Nil(R[x_0, ..., x_n]) \subseteq \a$ for all prime ideals $\a$, and the claim follows since $\Nil(R)$ is prime by assumption that $\Spec(R)$ is irreducible.
\end{proof}
\begin{prop}
    Assume $\Spec(R)$ is irreducible.
    Then $\P_R^n$ is an irreducible scheme.
\end{prop}
\begin{proof}
    We show that $\D{x_0} \subseteq \P_R^n$ is dense.
    The claim then follows, as any two nonempty, disjoint open sets give nonempty, disjoint open sets in $D_{x_0}$ which is isomorphic to the 
    irreducible\footnote{This is obviously irreducible, as $R[x_0, ..., x_n]/\Nil(R[x_0, ..., x_n]) = (R/\Nil(R))[x_0, ..., x_n]$ is integral by assumption and Lemma~\ref{prop:nil_poly_ring}.}
    affine scheme $\Spec(R[x_0, ..., x_n]_{x_0})_0$ by Lemma~\ref{prop:Df_affine}.

@@

    wlog $S$ is Noetherian, otherwise continue with the ring
    \begin{equation*}
        \tilde{S} := \bigl( \Z / \mathrm{char}(S) \Z \bigr) \bigl[ A/\alpha, x \bigr]
    \end{equation*}
    generated by the coefficients of $A/\alpha$ and $w$ (this is Noetherian, as it is a quotient of a polynomial ring in finitely many variables).

    Now equip $S_{\mathfrak{l}}$ with the $\mathfrak{l}$-adic topology.
    Note that $S$ is Noetherian and local, thus the Krull intersection theorem \cite[8.41]{comalg_notes} shows that the $\mathfrak{l}$-adic topology is Hausdorff and we find
    \begin{equation*}
        w = \lim_{i \to \infty} w = \lim_{i \to \infty} \frac {A^i} {\alpha^i} w = \Bigl( \lim_{i \to \infty} \frac {A^i} {\alpha^i} \Bigr) w = 0 w = 0
    \end{equation*}
    since $(A/\alpha)^i$ has coefficients in $\mathfrak{l}^i$, thus converges to $0$ as $i \to \infty$.

@@

But now enough of talking, here is the main corollary, which is also the main result of the miniproject.
\begin{corollary}
    The morphism $\P_R^n \to \Spec(R)$ is universally closed.
\end{corollary}
\begin{proof}
    Use Lemma~\ref{prop:universally_closed_affines}, so consider an affine base change $f_{\Spec(S)}: \P_R^n \times_{\Spec(R)} \Spec(S) \to \Spec(S)$.
    By Lemma~\ref{prop:projective_space_base_change}, have that $\P_R^n \times_{\Spec(R)} \Spec(S) \cong \P_S^n$ as schemes over $S$, hence $f_{\Spec(S)}$ is isomorphic to $\P_S^n \to \Spec(S)$.
    This morphism is closed by Proposition~\ref{prop:projective_structure_morphism_closed} and the claim follows.
\end{proof}
\begin{corollary}
    Projective space $\P_R^n$ over $\Spec(R)$ is proper.
\end{corollary}
\begin{corollary}
    Let $k$ be an algebraically closed field.
    Then projective space $\P_k^n$ over $\Spec(k)$ is a complete variety.
\end{corollary}

\section{Valuative criterion of properness}
\label{sec:valuative_criterion_properness}
Instead of using some magical ad-hoc argument as before, there is also the following nice characterization of the notion of properness.
We will omit the proof here, but present the statement as it can give more geometric intuition on properness and why $\P_R^n$ is proper.
\begin{prop}
    \label{prop:valuative_criterion_properness}
    Let $X$ be Noetherian and $f: X \to Y$ a morphism of finite type.
    The following are equivalent
    \begin{itemize}
        \item $f$ is proper
        \item For every valuation ring $R$ with field of fractions $K = \Frac(R)$ and all morphisms $\Spec(R) \to Y$, $\Spec(K) \to X$ that make the diagram
        \begin{center}
            \begin{tikzpicture}
                \node (K) at (0, 0) {$\Spec(K)$};
                \node (R) at (0, -1.5) {$\Spec(R)$};
                \node (X) at (3, 0) {$X$};
                \node (Y) at (3, -1.5) {$Y$};

                \draw [->] (K) -- (R);
                \draw [->] (K) -- (X);
                \draw [->] (R) -- (Y);
                \draw [->] (X) -- (Y) node [midway, right] {$f$};
            \end{tikzpicture}
        \end{center}
        commute, there exists a unique compatible morphism $\Spec(R) \to X$.
    \end{itemize}
\end{prop}
\begin{proof}
    Omitted, see e.g. \cite[II.4.7]{hartshorne} or \citestacks{0BX5}.
\end{proof}
\begin{prop}
    \label{prop:projective_space_proper_alternative_proof}
    The morphism $\P_R^n \to \Spec(R)$ is proper.
\end{prop}
\begin{proof}[Alternative Proof]
    Since properness is preserved under base change, by Lemma~\ref{prop:projective_space_base_change} it suffices to show the statement for $\P_\Z^n \to \Spec(\Z)$.

@@

            \draw [->] (K) -- (R);
            \draw [->] (K) -- (X);
            \draw [->] (R) -- (Y);
            \draw [->] (X) -- (Y) node [midway, right] {$f$};
        \end{tikzpicture}
    \end{center}
    commutes.
    Note that $\Spec(K) = \{ \langle 0 \rangle \}$, so let $\xi \in X$ be the image of $\langle 0 \rangle$ under $\Spec(K) \to X$.
    Since $\langle x_0, ..., x_n \rangle \not\subseteq \xi$, there is $x_i \not\in \xi$.
    Now let
    \begin{equation*}
        f_j := \begin{cases}
            x_j & \text{if $x_j \notin \xi_1$} \\
            x_j - x_i & \text{if $x_j \in \xi_1$}
        \end{cases}
    \end{equation*}
    Thus all $f_j$ are homogeneous of degree 1 with $\langle f_j \ | \ j \rangle = \langle x_0, ..., x_n \rangle$.
    It follows that $X = \bigcup_j \D{f_j}$ is a cover by affine opens such that $\xi \in \bigcap_j \D{f_j}$.
    Now possibly apply a linear automorphism $f_j \to x_j$, so have wlog $f_j = x_j$.
    In particular, find $x_j \in \O_{X, \xi_1}^*$.

    The morphism $\Spec(K) \to X$ now gives an inclusion $k(\xi) \subseteq K$.
    Since $x_j \in \O_{X, \xi}^*$, observe that $g_{ij} := x_i / x_j \in K$ is well-defined and nonzero with $g_{ij} g_{jk} = g_{ik}$.

    As $R$ is a valuation ring, we find $k$ such that $g_{k0}$ is the smallest element (w.r.t. divisibility by $R$) among $g_{00}, ..., g_{n0}$.
    So we see that $g_{ik} = g_{i0} / g_{k0} \in R$ for all $i$.
    In particular, we find a ring homomorphism
    \begin{equation*}
        \Z\left[ \frac {x_0} {x_k}, ..., \frac {x_n} {x_k} \right] \to R, \quad \frac {x_j} {x_k} \mapsto g_{i k}
    \end{equation*}
    that is compatible with the field inclusion $k(\xi) \subseteq K$.

@@

    It is not hard to show that the morphism is unique\footnote{In fact, the statement that there is at most one such morphism is equivalent to $\P_R^n \to \Spec(R)$ being separated (by \citestacks{01L0}), so follows from Proposition~\ref{prop:projective_space_separated}.}, and the claim follows.
\end{proof}
Note that in some way, this is the same proof as before.
The core idea - namely the choice of the smallest element $g_{k0}$ (w.r.t. divisibility by $R$) - more or less achieves the same effect as the more complicated ideal argument from Lemma~\ref{prop:adding_coeff_ideal_does_not_introduce_irrelevant_ideal}.
\printbibliography
\end{document}
